### 2022-11-14 project_Antonio/paper.tex
antonio proyecto

\def\Student{Antonio Paredes Sotelo}
\def\Title{Anteproyecto}
\def\Prog{Doctorado en Ciencias (F\'{i}sica) }
\def\Dept{Departamento de Investigac\'{i}on en Fis\'{i}ca}
\def\Director{Dr. Jos\'{e} Feliciano Ben\'{i}tez Rubio}
\def\ProjectTitle{B\'usqueda de part\'iculas de Higgs con modelos de f\'sica m\'as all\'a del Modelo Est\'andar y medici\'on de la luminosidad usando el m\'etodo de Pixel Cluster Counting en el experimento CMS del CERN}
\def\ProjectTitleEnglish{Search for Higgs bosons using models of physics beyond the Standard Model and measurement of luminosity using the Pixel Cluster Counting method in the CMS experiment at CERN}
\def\ResearchLine{F\'{i}sica Matematicas}

@@

%%%% Abstract Page
\newpage
\hspace{2pt}
\vfill

@@

\iffalse
\onehalfspacing This research project proposes to study the unobserved production channel of the Higgs boson in associated with a single top quark (tH) using Higgs boson decay to multilepton final state. The dataset collected during the Run II (2015-2018) of Large Hadron Collider (LHC) will be used corresponding to an integrated luminosity of 150 $fb^{-1}$. Increasing the collider luminosity is expected to increase the number of event for this Higgs boson production mode. The measurement of this production mode has ramifications for other measurements related to coupling parameter $y_t$ of the top quark and the Higgs boson which is part of the physics program of the international CMS and ATLAS collaborations. Previous measurements using dataset with an integrated luminosity 36 $fb^{-1}$ has put an upper limit of 2.04 pb on the production cross section which is far from standard model predicted value of 0.077 pb. We plan to obtain more stringent upper limits on the production cross section and get constraints on the Higgs top quark coupling parameter $y_t$. \par
\fi

\onehalfspacing This research project proposes to study the unobserved decays modes of the Higgs boson. \iffalse associated with a single top quark (tH) using Higgs boson decay to multilepton final state \fi The dataset collected during the Run II (2015-2018) of Large Hadron Collider (LHC) will be used corresponding to an integrated luminosity of 150 $fb^{-1}$. Serches for decays of the Higgs boson into pairs of $J/\psi$ or $\Upsilon$.

%Increasing the collider luminosity is expected to increase the number of event for this Higgs boson production mode. The measurement of this production mode has ramifications for other measurements related to coupling parameter $y_t$ of the top quark and the Higgs boson which is part of the physics program of the international CMS and ATLAS collaborations. Previous measurements using dataset with an integrated luminosity 36 $fb^{-1}$ has put an upper limit of 2.04 pb on the production cross section which is far from standard model predicted value of 0.077 pb. We plan to obtain more stringent upper limits on the production cross section and get constraints on the Higgs top quark coupling parameter $y_t$. \par

%\onehalfspacing In this project, it is proposed to search for the decays of the Higgs boson to pairs of $J/\psi$ or $upsilon$ in the CMS detector based at the Large Hadron Collider of the CERN laboratory.

%For these studies the data of the LHC Run 2

@@

\onehalfspacing The standard model (SM) of particle physics is so far the best theoretical model to describe the interaction of elementary 
particles using three of the four fundamental forces of nature which are electromagnetic force, strong nuclear force and the 
weak nuclear force. Gravitational force is neglected as the strength of this force is very weak at the scales over which 
elementary particle interact with each other. The standard model (SM) of particle physics is divided into two categories, 
bosonic sector and fermionic sector. Bosonic sector contain particles called bosons which mediate the fundamental forces of 
nature and the fermionic sector contain particles called fermions which make up all the matter in our universe. SM has three 
generations of matter (fermions) particles. The first generation of fermions consists of up (u) quark, down (d) quark, electron 
and electron neutrino, second generation consist of charm (c) quark, strange (s) quark, muon and muon neutrino and the third 
generation of matter particles has top (t) quark, bottom (b) quark, tau and tau neutrino. The bosonic sector consist of gauge 
bosons like gluon, photon, $W^{\pm}$, $Z^0$ which mediate strong nuclear force, electromagnetic force and weak nuclear force respectively. There is one more particle in the standard model called the Higgs Boson which gives mass to SM particles via electroweak symmetry breaking mechanism \cite{Chatrchyan:2012xdj}. All the standard model particles are shown in Figure 1. Higgs boson can be produced at the particle colliders like the Large Hadron Collider (LHC) in Geneva, Switzerland.

\begin{figure}[H]
	\centering
	\includegraphics[width= 0.5 \columnwidth]{./sm.png}
	\caption{Elementary particles in standard model of particle physics with their mass, spin and charge.}
	\label{figure 1}
\end{figure}

% \newpage

There are various production modes of Higgs boson like the gluon gluon fusion (ggF), vector boson fusion (VBF), Higgs 
production in association with vector boson (VH, V = $W^{\pm}$ or Z), Higgs production with top quark and top anti-quark ($t\bar{t}H$) and Higgs 
production in association with a single top quark and a quark jet (tHq). Each production channel has its own importance to probe the 
various properties like spin, parity and the coupling strength parameters of Higgs boson to fermions and bosons. The main production mode of Higgs boson at LHC is 
gluon gluon fusion (ggF). Higgs boson can decay to a pair of $W^{\pm}$ bosons, Z bosons, photons, bottom quarks, 
muons and electrons. Higgs boson can also decay to Z boson and a photon. The important SM Higgs Boson production modes are shown in Figure 2. The production cross section and branching ratios of various decay channels of Higgs boson are shown in Figure 3.
Figure 3 also show the main decays channels by the SM.

%\afterpage{
\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{./production_mode.png}
	\caption{Important production processes of Standard model Higgs boson ggF, VBF, VH in proton collisions.}
	\label{figure 2}
\end{figure}

\begin{figure}[H]
  \centering
   \includegraphics[scale=0.7]{./cd2.png}
  \caption{Standard model Higgs boson production cross section with center of mass energy and branching ratios for various decay channels \cite{Tanabashi:2018oca}.}
   \label{figure 3}
\end{figure}
\clearpage

@@

%\afterpage{
  \begin{figure}[H]
    \centering
    \includegraphics[width=0.6\columnwidth]{./cms12.png}
    \caption{Transverse view of the CMS detector showing the silicon tracker, electromagnetic calorimeter, hadron calorimeter, superconducting solenoid and muon chambers \cite{Chatrchyan:2008aa}.}
    \label{figure5}
  \end{figure}

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\columnwidth]{./lum6.png}
    \caption{Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS) of the LHC and luminosities. Points show instantaneous luminosity while the line shows luminosity accumulated \cite{collaborations2019report}.}
    \label{figure6}
  \end{figure}

%\clearpage

@@

\clearpage
\iffalse
The LHC will deliver about 300 $fb^{-1}$ by 2024 \cite{collaborations2019report}.
The goal of subsequently run of LHC has been to obtain datasets at higher values of luminosities for precision measurements of the properties of Higgs boson, in order to test the Standard Model pattern of couplings to elementary particles.
In order to minimize the machine downtimes and maximize the productive use of the LHC for physics, the replacement of the inner triplet magnets (the one responsible to squeeze the beam at collision) and  of  all  hardware  changes  needed  to  enable  an  ambitious  luminosity  upgrade  will  take  place in parallel during one shutdown, at around 2023-25 (LS3), with some of the modification anticipated in 2019-2020 (LS2).

This new phase of the LHC life has been named as High Luminosity LHC (HL-LHC) and has the scope of attaining the astonishing threshold of 3000 $fb^{-1}$ by the year 2036 as shown in Figure~\ref{figure6}.
Figure~\ref{figureKappas} shows a summary of the current measurements of the Higgs coupling parameters with the LHC Run 1 dataset (2011,2012)  using ggF, VBF, VH, and ${t\bar{t}H}$ production modes \cite{Tanabashi:2018oca}.
Higgs boson production in association with single top quark has been studied to derive constraints on the magnitude and relative sign of Higgs boson couplings to top quarks and vector bosons.

Constaints on $\kappa_t$ can be derived using a likelihood ratio scan of $L(\kappa_t)/L(\hat{\kappa_t})$ where $\hat{\kappa_t}$ is the best fit value of $\kappa_t$ as shown in Figure~\ref{figureKt} \cite{Sirunyan:2018lzm}.
Multilepton final states like ee, $\mu\mu$ and $e\mu$ and final states with single lepton with a pair of bottom quarks were combined with Higgs decay to two photons to get the final result. The data slightly favor positive value of top quark yukawa coupling for SM Higgs boson and excludes ranges outside the intervals [-0.9,-0.5] and [1.0,2.1] times $y^{SM}_t$ at the 95 $\%$ confidence level.

\newpage
%\afterpage{
  \begin{figure}[H]
    \centering
    \includegraphics[scale=0.4]{./couplings.png}
    \caption{ATLAS-CMS combined measurements of coupling strength modifiers \cite{Tanabashi:2018oca}.}
    \label{figureKappas}
  \end{figure}

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.4 \columnwidth]{./cms13.png}
    \caption{Scan of -2 $\Delta$ ln(L) versus $\kappa_t$ for the 2016 LHC data (black line) and the individual channels (blue, red, and green), compared to Asimov data sets corresponding to the SM expectations (dashed lines) \cite{sirunyan2019search}.}
    \label{figureKt}
  \end{figure}

\newpage
\fi
\section{PROPOSAL}

\iffalse
\onehalfspacing In this project, it is proposed to search for the production of Higgs boson associated with a single top quark in the CMS detector based at the Large Hadron Collider of the CERN laboratory.
It is proposed to study the sensitivity with the current data in the channel where the events are identified with two muons of the same sign, electron muon or with three leptons of electron or muon type.
Leptons of tau type are not considered due to complications in their identification in the detector.
\fi

\onehalfspacing In this project, it is proposed to search for the decays of the Higgs boson to pairs of $J/\psi$ or $\Upsilon$ in the CMS detector based at the Large Hadron Collider of the CERN laboratory. For these studies the data of the LHC Run 2 (2015-2018) will be used corresponding to an integrated luminosity of about 150 fb$^{-1}$, this is 4 times larger than the dataset used for current results. Previously CMS has carried out studies with the data from 2017 ~\cite{2019134811} corresponding to 37.5 fb$^{-1}$ and ATLAS in 2015 ~\cite{PhysRevLett.114.121801} corresponding to 20.3 fb$^{-1}$.

%It is also expected to estimate projections of the sensitivity to the High Luminosity phase of the LHC (HL-LHC) (2026-2036) which will accumulate up to 100 times the data used for current measurements.

%It is proposed to study the sensitivity with the current data in the channel where the events are identified with two muons of the same sign, electron muon or with three leptons of electron or muon type. Leptons of tau type are not considered due to complications in their identification in the detector.

\section{GENERAL OBJECTIVE}

\onehalfspacing There is currently a great enigma in physics since we know that ordinary matter comprises only $5\%$ of the universe, another $27\%$ includes Dark matter and the rest $68\%$ is Dark energy.
Understanding the production of the Higgs boson, as well as its decays are an important part of the physics program of CERN laboratory experiments to verify the Standard Model and look for new particles. Through this project we will investigate the decay of the Higgs boson to pairs of $J/\psi$ or $\Upsilon$ in proton-proton collisions with the CMS experiment at the LHC. An observation of such decay  with this sample would indicate the presence of physics beyond the standard model. The decay to pairs of $J/\psi$ or $\Upsilon$ has not been observed so far.

%An observation of such decay  with this sample would indicate the presence of physics beyond the standard model. 
\iffalse
\onehalfspacing There is currently a great enigma in physics since we know that ordinary matter comprises only $5\%$ of the universe, another $27\%$ includes Dark matter and the rest $68\%$ is Dark energy.
Understanding the production of the Higgs boson, as well as its decays are an important part of the physics program of CERN laboratory experiments to verify the Standard Model and look for new particles. Through this project we will investigate the production of the Higgs boson in association with a single top quark (tH) in proton-proton collisions with the CMS experiment at the LHC. The tH  production process has not been observed so far.
\fi

\iffalse
The measurement of this process complements other measurements of the yukawa coupling parameter $y_t$ of the top quark and Higgs boson and is a part of the physics program of CMS and ATLAS international collaborations. Through these measurements we seek to complete the measurements of the Standard Model parameters to high precision and also to find evidence of deviations in yukawa couplings which point to new particles which may explain current enigmas like the origin of Dark Matter.
\fi

\section{HYPOTHESIS}
%%  - what we expect to find

% Studies about the decay of the Higgs boson have already started within the CMS and ATLAS collaboration and publications are based on data from the 2011-2012 run and some data from the second run of 2015-2016.

\onehalfspacing With the CMS search mentioned above using 2017 data, an upper limit of $1.8 \times 10^{-3}$ was obtained in the $J/\psi$ pair while in the $\Upsilon$ pair channel, the limit was $1.4 \times 10^{-3}$. With this project we propose to incorporate the full Run 2 dataset corresponding to 150 fb$^{-1}$. A factor of 4 increase in number of collision events will improve the limits by a factor of approximately 2 due to the nature of statistical uncertainties.

%In case of observing a signal with this data this would mean evidence of physics beyond the standard model.
\iffalse
Additionally, in subsequent phases of the LHC (2021-2023 and 2026-2038) the amount of data will increase by a factor of up to 100, which implies the upper limit for the SM case will reduce to about 2.6 times the expected cross-section. This implies that even for the HL-LHC it will be difficult to observe the tH production, but at the same time more and more phase space of BSM models will be excluded.

%these were the results
%The currently published result \cite{} uses 37.5 fb$^{-1}$ of collision data and obtains 
%CMS collaboration has perfomed
%pairexpected in the SM \cite{sirunyan2019search}, the limit is 26 times larger than the expected value.
%The most sensitive channel is  with leptons, however this channel is a combination of two sub-channels: two-lepton and three-lepton.
%The currently published result uses 36 fb$^{-1}$ of collision data and obtains an upper limit of 2.04 pb while only 0.077 pb are expected in the SM \cite{sirunyan2019search}, the limit is 26 times larger than the expected value.
%In BSM models, for example with $\kappa_t = -1$,  the expected cross-section increases up to 0.834 pb due to quantum interference effects; in this case also the observed upper limit changes to 0.86 pb at 95$\%$ confidence level, the current data almost rules out this model.

%This mean that the limit for the SM case will improve to about 13 times the expected cross-section, while for the modified model ($\kappa_t = -1$) we expect to be excluded.
In this search the reconstruction methods do not allow to separate $t$H from $t\bar{t}H$ production because the cross-section of $t\bar{t}H$ is much larger than that of tH and some events pass the selections designed for tH.
The fraction of tH events over tH + $t\bar{t}H$ is currently only 5$\%$ with the analysis techniques used.
We expect to improve the analysis techniques using multi-variate techniques, for example boosted decision trees or other algorithms, to improve the identification of signal events over the rest of the backgrounds.
This will provide additional improvements in the limits and a possible observation of the SM signal as well as stringent exclusions for BSM models.
\fi

%\newpage
\section{SPECIFIC OBJECTIVES}

\onehalfspacing

\iffalse
The measurement of the tH production mechanism is a collaborative effort in the LHC experiment due to the various channels that need to be combined.
This project focuses on the channels with leptons in the final state: two-lepton ($\ell\ell$) or three-lepton ($\ell\ell\ell$).
The $\ell\ell$ sub-channel uses events with same sign dimuon ($\mu^\pm\mu^\pm$) or electron-muon ($e^\pm\mu^\pm$).
The $\ell\ell\ell$ sub-channel uses events in the topologies: $\mu\mu\mu$, $\mu\mu e$, $\mu e e$ and $e e e$, where no charge requirements are made.
The leptons are generated from the decays of the $W$ and $Z$ bosons or $\tau$ leptons from decays of the Higgs or top quarks.
Furthermore, light-quark and b-quark jets are generated from the initial state collision and from the decay of the top quark, these must be used in the signal identification.
\fi

In order to complete a journal publication of this search the following are need it:
%production using one of these channels, several aspects  must be worked on:
%of these search
\begin{itemize}
\item Simulation of the signal events using Monte Carlo generators which implement the theoretical model.
\item Implementation of event reconstruction and selection algorithms which select high quality signal events and reduce the background events to a minimum.
\item Studies of lepton algorithms which improve the background rejection.
\item Due to the small signal over background ratio, a multi-variate algorithm is maybe employed which combines different event variables and exploits the correlations to construct a final discriminant. This algorithm must be trained using simulated signal and background events.
\item A model for the observed data must be constructed from the known background processes. 
%These processes are normally simulated individually and then summed together with the signal to construct the final statistical model.
\item Reconstruction of $J/\psi$ and $\Upsilon$ with two muons ~\cite{Khachatryan_2014,Sirunyan_2018}.
\item An algorithm is developed that adjusts the model to the observed data distribution and incorporates several systematic uncertainties due to differences between simulation and real data.
\item The signal strength observed from the statistical analysis must be interpreted as evidence of a signal or in the absence as a limit. %Further analysis is required to place constraints on paramater space of BSM models.
\end{itemize}

@@

%During regular data-taking periods, activities include monitoring the data acquisition systems and data quality through comparisons of new data and old data.
%In addition to the data analysis plans described before, another objective is to contribute in one of the developments of the CMS detector to  prepare for the HL-LHC phase.
%

%Finally, the results of this project will be presented at national or international conferences and a publication in a scientific journal will be accomplished.

\section{METHODOLOGY}
%- methods used in the search: which dataset,  lepton reconstruction, selections, jets, BDT, backgrounds, statistical analysis

%\afterpage{
%\begin{figure}[H]
%	\centering
%	\includegraphics[width=0.5\columnwidth]{./topology.png}
%	\caption{Topology of a tH signal event in a proton-proton collision. The decay products from the top quark include a b-jet and a W boson decaying to a lepton and neutrino. In this diagram the Higgs decays to a pair of W bosons, one decays to a lepton and neutrino while the other one can decay leptonically or hadronically. In addition there is a forward jet from initial collision of the quarks.}
%	\label{figureTopology}
%\end{figure}
%\clearpage

%\onehalfspacing The methods for a measurement of the 
%coupling of the Higgs boson and the top quark in proton collisions are extensive and require several stages.

@@

\iffalse
\begin{figure}[H]
	\centering
	\includegraphics[width=0.5\columnwidth]{./topology.png}
	\caption{Topology of a tH signal event in a proton-proton collision. The decay products from the top quark include a b-jet and a $W^{\pm}$ boson decaying to a lepton and neutrino. In this diagram the Higgs decays to a pair of $W^{\pm}$ bosons, one decays to a lepton and neutrino while the other one can decay leptonically or hadronically. In addition there is a forward jet from initial collision of the quarks.}
	\label{figureTopology}
\end{figure}
\fi

%Figure~\ref{figureTopology} shows the main features of the topology of a tH event, including the subsequent decay products of the top quark and Higgs boson, these are the particles which are generated at the interaction point before entering the detector components.
A separate simulation of the detector is employed using \textsc{Geant}~\cite{agostinelli2003geant4} which simulates the trajectory of the stable particles, response and efficiency of each sensor in the CMS detector.
The information obtained from this chain is saved as raw data similar to the one obtained from the real experiment and is later reconstructed in the same way.

The event reconstruction is designed to capture the distinguishing features of the signal topology.
%In the channels which are the focus of this project, we require that there exist at least two leptons of electron or muon type, however also jets need to be identified as shown in the Figure 9.
The event reconstruction begins with the identification of the different types of particles: photon, electrons, muons, taus, and jets.
Particle lists are formed for each kind using a general purpose event reconstruction software called \textsc{CMSSW}~\cite{Bayatian:922757} which is maintained by the collaboration and which includes basic calibration of the various particle energy and momenta.

@@

%Jets which originate from b quarks are important in this study because they identify the decay of a top quark, b-jets are identified with a multivariate algorithm which searches for secondary vertices in the Tracker and electrons or muons (from decays of B-mesons) within the cone of the jet.

Luminosity is one of the most omportant parameter for a particle accelerator. It is the measure of number of proton-proton collisions occurring in unit centimetre square area per second. The installation of the CMS Phase-1 pixel detector took place during the extended year-end technical stop of the LHC in 2016/2017. The CMS Phase-1 pixel detector consists of four concentric barrel layers anf three disks. The CMS Phase-1 pixel detector is built from 1856 segmented silicon sensor modules,
where 1184 modules are used in the barrel pixel detector (BPIX) and 672 modules are used for the forward disks (FPIX). Each module consists of a sensor with 160x416 pixels connected to 16 readout chips (ROCs).One of the systems used for measuring luminosity at CMS is a technique called Pixel Cluster Counting (PCC) method. The PCC method uses the rate of pixel clusters in the CMS pixel detector to provide an luminosity measurement.

@@

%In the search for tH events using the dimuon channel the following event selections are applied:
\iffalse
\begin{itemize}
\item
\item
\item
\item

\end{itemize}
\fi

%\textsc{Root} is a \textsc{c++} based software that allows visualization of event distributions and variable correlations. Further optimization of the event selections is performed by considering the different processes which contribute to the background. In this search several SM processes contribute to the background due to their decays to 
%leptons: $pp\rightarrow ZZ, WZ, WW, ZZZZ, t\bar{t}W, t\bar{t}Z, t\bar{t}H, tWZ, tZq,$ and $t\bar{t}t\bar{t}$. These backgrounds must be determined from MC simulation and normalized to their predicted SM cross sections.

@@

\section{EXPECTED RESULTS}
%- a limit, also predictions for future runs

\onehalfspacing The results from this project include the following items:
\begin{itemize}
\item The student will become part of an international collaboration.
\item A leading contribution in one of the search for decays of the Higgs boson into pairs of $J/\psi$ or $\Upsilon$ using the data from the Run II of the LHC, resulting in a publication in a scientific journal.
%\item Participation in the CMS detector Phase II upgrade through development and possible installation of particle detectors possibly resulting in a technical report.
\item Improve limits on the branching fraction of these decays channels.
\item Luminosity study for RUN3 data.
\item A presentation of a poster or a talk in one of the national conferences of the Division of Paticles and Fields of the Mexican Society of Physics or a presentation of a poster or a talk in an international physics conference.
\item An academic stay at a scientific center like CERN or Fermilab.
\item A high quality thesis.
\end{itemize}

>>>>

\def\Student{Antonio Paredes Sotelo}
\def\Title{Anteproyecto}
\def\Prog{Doctorado en Ciencias (F\'{i}sica) }
\def\Dept{Departamento de Investigac\'{i}on en Fis\'{i}ca}
\def\Director{Dr. Jos\'{e} Feliciano BEN\'{I}TEZ RUBIO}
\def\ProjectTitle{B\'usqueda de part\'iculas de Higgs con modelos de f\'sica m\'as all\'a del Modelo Est\'andar y medici\'on de la luminosidad usando el m\'etodo de Pixel Cluster Counting en el experimento CMS del CERN}
\def\ProjectTitleEnglish{Search for Higgs bosons using models of physics beyond the Standard Model and measurement of luminosity using the Pixel Cluster Counting method in the CMS experiment at CERN}
\def\ResearchLine{F\'{i}sica Matematicas}

@@

%%%% Abstract Page
\newpage
\hspace{2pt}

@@

\onehalfspacing This research project proposes to study the unobserved decays modes of the Higgs boson into pairs of $J/\psi$ or $\Upsilon$.  The dataset collected during the Run II (2015-2018) of Large Hadron Collider (LHC) will be used corresponding to an integrated luminosity of 150 $fb^{-1}$. Previous searches for these decays were done with only 37 $fb^{-1}$, the increased luminosity and improved analysis strategy will provide more stringent limits for these decays and more sensitivity to physics beyond the Standard Model.

@@

\onehalfspacing The Standard Model (SM) of particle physics is so far the best theoretical model to describe the interaction of elementary 
particles using three of the four fundamental forces of nature which are electromagnetic force, strong nuclear force and the 
weak nuclear force. Gravitational force is neglected as the strength of this force is very weak at the scales over which 
elementary particle interact with each other. The standard model (SM) of particle physics is divided into two categories, 
bosonic sector and fermionic sector. Bosonic sector contain particles called bosons which mediate the fundamental forces of 
nature and the fermionic sector contain particles which make up all the matter in our universe. SM has three 
generations of matter (fermions) particles. The first generation of fermions consists of up (u) quark, down (d) quark, electron 
and electron neutrino, second generation consist of charm (c) quark, strange (s) quark, muon and muon neutrino and the third 
generation of matter particles has top (t) quark, bottom (b) quark, tau and tau neutrino. The bosonic sector consist of gauge 
bosons like gluon, photon, $W^{\pm}$, $Z^0$ which mediate strong nuclear force, electromagnetic force and weak nuclear force respectively.
There is one more particle in the standard model called the Higgs boson which gives mass to SM particles via electroweak symmetry breaking mechanism \cite{Chatrchyan:2012xdj}. All the standard model particles are shown in Figure 1. Higgs boson can be produced at the particle colliders like the Large Hadron Collider (LHC) in Geneva, Switzerland.

There are various production modes of Higgs boson like the gluon gluon fusion (ggF), vector boson fusion (VBF), Higgs 
production in association with vector boson (VH, V = $W^{\pm}$ or Z), Higgs production with top quark and top anti-quark ($t\bar{t}H$) and Higgs 
production in association with a single top quark and a quark jet (tHq). Each production channel has its own importance to probe the 
various properties like spin, parity and the coupling strength parameters of Higgs boson to fermions and bosons.
The main production mode of Higgs boson at LHC is gluon gluon fusion (ggF).
Higgs boson can decay to a pair of $W^{\pm}$ bosons, Z bosons, photons, bottom , charm, or strange quarks, also to taus and muons .
The important SM Higgs Boson production modes are shown in Figure 2. The production cross section and branching ratios of various decay channels of Higgs boson are shown in Figure 3.
Figure 3 also show the main decays channels by the SM.

\clearpage
\begin{figure}[H]
	\centering
	\includegraphics[width= 0.5 \columnwidth]{./sm.png}
	\caption{Elementary particles in standard model of particle physics with their mass, spin and charge.}
	\label{figure 1}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{./production_mode.png}
	\caption{Important production processes of Standard model Higgs boson ggF, VBF, VH in proton collisions.}
	\label{figure 2}
\end{figure}

\begin{figure}[H]
  \centering
   \includegraphics[scale=0.7]{./cd2.png}
  \caption{Standard model Higgs boson production cross section with center of mass energy and branching ratios for various decay channels \cite{Tanabashi:2018oca}.}
   \label{figure 3}
\end{figure}

\clearpage

@@

@@

\clearpage
  \begin{figure}[H]
    \centering
    \includegraphics[width=0.6\columnwidth]{./cms12.png}
    \caption{Transverse view of the CMS detector showing the silicon tracker, electromagnetic calorimeter, hadron calorimeter, superconducting solenoid and muon chambers \cite{Chatrchyan:2008aa}.}
    \label{figure5}
  \end{figure}

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\columnwidth]{./HLLHCLumi.png}
    \caption{Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS) of the LHC and luminosities. Points show instantaneous luminosity while the line shows luminosity accumulated \cite{collaborations2019report}.}
    \label{figure6}
  \end{figure}

\clearpage
\section{PROPOSAL}
\onehalfspacing In this project, it is proposed to search for the decays of the Higgs boson to pairs of $J/\psi$ or $\Upsilon$ in the CMS detector based at the Large Hadron Collider of the CERN laboratory. For these studies the data of the LHC Run 2 (2015-2018) will be used corresponding to an integrated luminosity of about 150 fb$^{-1}$, this is 4 times larger than the dataset used for current results. Previously CMS has carried out studies with the data from 2017 ~\cite{2019134811} corresponding to 37.5 fb$^{-1}$ and ATLAS in 2012 ~\cite{PhysRevLett.114.121801} corresponding to 20.3 fb$^{-1}$.

\onehalfspacing We also propose to work on the luminosity measurement for the new LHC Run 3 which has just started in the year 2022. The measurement of the integrated luminosity is essential for many physics studies at CMS.

\section{GENERAL OBJECTIVE}
\onehalfspacing There is currently a great enigma in physics since we know that ordinary matter comprises only $5\%$ of the universe, another $27\%$ includes Dark matter and the rest $68\%$ is Dark energy.
Understanding the production of the Higgs boson, as well as its decays are an important part of the physics program of CERN laboratory experiments to verify the Standard Model and look for new particles. Through this project we will investigate the decay of the Higgs boson to pairs of $J/\psi$ or $\Upsilon$ in proton-proton collisions with the CMS experiment at the LHC. An observation of such decay  with this sample would indicate the presence of physics beyond the standard model. The decay to pairs of $J/\psi$ or $\Upsilon$ is expected to be extremely small, below  current detection limitations.

\section{HYPOTHESIS}
\onehalfspacing With the CMS search mentioned above using 2017 data, an upper limit of $1.8 \times 10^{-3}$ was obtained in the $J/\psi$ pair while in the $\Upsilon$ pair channel, the limit was $1.4 \times 10^{-3}$. With this project we propose to incorporate the full Run 2 dataset corresponding to 150 fb$^{-1}$. A factor of 4 increase in number of collision events will improve the limits by a factor of approximately 2 due to the nature of statistical uncertainties.

\section{SPECIFIC OBJECTIVES}

\onehalfspacing 
In order to complete a journal publication of this search the following are needed:
\begin{itemize}
\item Simulation of the signal events using Monte Carlo generators which implement the theoretical model.
\item Implementation of event reconstruction and selection algorithms which select high quality signal events and reduce the background events to a minimum.
\item Studies of lepton algorithms which improve the background rejection.
\item Reconstruction of $J/\psi$ and $\Upsilon$ with two muons ~\cite{Khachatryan_2014,Sirunyan_2018}.
\item Due to the small signal over background ratio, a multi-variate algorithm may be employed which combines different event variables and exploits the correlations to construct a final discriminant. This algorithm must be trained using simulated signal and background events.
\item A model for the observed data must be constructed from the known background processes. 
\item An fitting algorithm is developed that adjusts the model to the observed data distribution and incorporates several systematic uncertainties due to differences between simulation and real data. The signal strength observed from the statistical analysis must be interpreted as evidence of a signal or in the absence as a limit. 
\end{itemize}

@@

\section{METHODOLOGY}

@@

A separate simulation of the detector is employed using \textsc{Geant}~\cite{agostinelli2003geant4} which simulates the trajectory of the stable particles, response and efficiency of each sensor in the CMS detector.
The information obtained from this chain is saved as raw data similar to the one obtained from the real experiment and is later reconstructed in the same way.

The event reconstruction is designed to capture the distinguishing features of the signal topology.
The event reconstruction begins with the identification of the different types of particles: photon, electrons, muons, taus, and jets.
Particle lists are formed for each kind using a general purpose event reconstruction software called \textsc{CMSSW}~\cite{Bayatian:922757} which is maintained by the collaboration and which includes basic calibration of the various particle energy and momenta.

@@

Luminosity is one of the most important parameter for a particle accelerator. It is the measure of number of proton-proton collisions occurring in unit centimetre square area per second. The installation of the CMS Phase-1 pixel detector took place during the extended year-end technical stop of the LHC in 2016/2017. The CMS Phase-1 pixel detector consists of four concentric barrel layers anf three disks. The CMS Phase-1 pixel detector is built from 1856 segmented silicon sensor modules,
where 1184 modules are used in the barrel pixel detector (BPIX) and 672 modules are used for the forward disks (FPIX). Each module consists of a sensor with 160x416 pixels connected to 16 readout chips (ROCs).One of the systems used for measuring luminosity at CMS is a technique called Pixel Cluster Counting (PCC) method. The PCC method uses the rate of pixel clusters in the CMS pixel detector to provide an luminosity measurement.

@@

@@

\section{EXPECTED RESULTS}
\onehalfspacing The results from this project include the following items:
\begin{itemize}
\item Becoming author of an international collaboration.
\item A leading contribution in one of the search for decays of the Higgs boson into pairs of $J/\psi$ or $\Upsilon$ using the data from the Run II of the LHC, resulting in a publication in a scientific journal.
\item Improve limits on the branching fraction of these decays channels.
\item Luminosity measurements for RUN3 data.
\item A presentation of a poster or a talk in one of the national conferences of the Division of Paticles and Fields of the Mexican Society of Physics or a presentation of a poster or a talk in an international physics conference.
\item An academic stay at a scientific center like CERN or Fermilab.
\end{itemize}

\cleardoublepage
\section{CALENDAR OF ACTIVITIES}
\onehalfspacing
\begin{itemize}
\setlength\itemsep{-0.5em}
\item {\bf Semester 1 (2022-2)}: 
  \SubItem{ Readings on Standard Model theory}
  \SubItem{ Basic Linux computing skills (\textsc{Bash, Emacs, Root})}
  \SubItem{ Initial planning of the analysis strategy}
  \SubItem{ Initial work on 2022 luminosity measurement}
\item {\bf Semester 2 (2023-1)}:
  \SubItem{ Course I on particle physics and/or particle detection and analysis}
  \SubItem{ Basic Linux computing skills (\textsc{Bash, Emacs, Root})}
  \SubItem{ Readings on SM and $J/\psi$ and $\Upsilon$ literature}
  \SubItem{ Computing accounts at CERN and Fermilab}
  \SubItem{ Studies of the $J/\psi$ and $\Upsilon$ event selections}
  \SubItem{ Work on 2022 luminosity measurement}
  \SubItem{ Possible summer stay at CERN or Fermilab}
\item {\bf Semester 3 (2023-2)}:
  \SubItem{ Course II on particle physics and/or particle detection and analysis}
  \SubItem{ Studies of the $J/\psi$ and $\Upsilon$ event selections}
  \SubItem{ Studies of the background processes}
  \SubItem{ Work on 2023 luminosity measurement}
\item {\bf Semester 4 (2024-1)}:
  \SubItem{ Studies of the $J/\psi$ and $\Upsilon$ event selections}
  \SubItem{ Development of a Multivariate algorithm for the event selections}
  \SubItem{ Studies of the $J/\psi$ and $\Upsilon$ background processes}
  \SubItem{ Development of the statistical model and signal extraction}
  \Subitem{ Work on 2023 luminosity measurement}
  \SubItem{ Possible summer stay at CERN or Fermilab}
\item {\bf Semester 5 (2024-2)}:
  \SubItem{ Studies of the $J/\psi$ and $\Upsilon$ event selections}
  \SubItem{ Development of a Multivariate algorithm for the event selections}
  \SubItem{ Studies of the $J/\psi$ and $\Upsilon$ background processes}
  \SubItem{ Development of the statistical model and signal extraction}
  \Subitem{ Work on 2024 luminosity measurement}
\item {\bf Semester 6 (2025-1)}:
  \SubItem{ Completion of the Higgs reconstruction and background estimation methods}
  \SubItem{ Development of software for the statistical analysis and interpretation of the results}
  \SubItem{ Review process of the results within the CMS collaboration}
\item {\bf Semester 7 (2025-2)}:
  \SubItem{ Completion of the analisis work}
  \SubItem{ Writing of the paper publication}
  \SubItem{ Presentation of poster or talk at national or international conference}
\item {\bf Semester 8 (2026-1)}:
  \SubItem{ Writing of the thesis}
  \SubItem{ Completion of the paper publication}

\end{itemize}
